---
title: "AmNatHistoryModels"
author: "FD"
output: 
    html_document: 
          toc: true
          toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Data

## Data sources

On 2020-07-20, I searched for all articles with `Publication Name = American Naturalist` on Web of Science Core collection. There were 11029 results. I manually exported all results in batches of 500 (export limit on WoS, when you also want to export the abstract), as `Tab-delimited (Mac)`. I collated the downloaded files, opened the merged file in Excel; I removed empty columns, renamed columns, as saved the file as `.csv`.

The source files are in <https://github.com/flodebarre/2020_AmNatHistory/tree/master/data/sourcesWOS/withAbstract>. 

## Load data 

```{r}
allArticles <- read.csv("../data/AmNat_allAbstracts.csv", stringsAsFactors = FALSE)
```

## Explore data

Number of articles loaded: `r nrow(allArticles)` items in the dataset.

```{r }
library(RColorBrewer)
cols <- brewer.pal(6, "Set2")
maincol <- cols[1]
```

Number of articles per year:
```{r}
nb.byY <- aggregate(allArticles$Authors, by = list(allArticles$PublicationYear), FUN = length)
par(las = 1)
plot(nb.byY$Group.1, nb.byY$x, 
     xlab = "Publication Year", ylab = "Number of items", 
     ylim = c(0, max(nb.byY$x)), axes = FALSE, 
     pch = 16, col = maincol)
axis(1, pos = 0)
axis(2, pos = 1900-1)
```

Articles with abstract:

```{r}
# Add information about number of characters in the abstract of each article
allArticles <- cbind(allArticles, AbsLength = nchar(allArticles$Abstract))

# Compute proportion of articles with non zero abstract length, by year
isAbs.byY <- aggregate((!is.na(allArticles$Abstract)), by = list(allArticles$PublicationYear), FUN = mean)

# Plot the result
par(las = 1)
plot(isAbs.byY$Group.1, isAbs.byY$x, 
     xlab = "Publication Year", ylab = "Proportion of items with non zero abstract length", 
     pch = 16, col = cols[2], 
     axes = FALSE)
axis(1, pos = 0)
axis(2, pos = 1900-1)
```

Sample sizes, items with abstracts, per year
```{r}
# Compute proportion of articles with non zero abstract length, by year
isAbs.byY <- aggregate((!is.na(allArticles$Abstract)), by = list(allArticles$PublicationYear), FUN = sum)

# Plot the result
par(las = 1)
plot(isAbs.byY$Group.1, isAbs.byY$x, 
     xlab = "Publication Year", ylab = "Number of items with non zero abstract length", 
     pch = 16, col = cols[2], 
     axes = FALSE)
axis(1, pos = 0)
axis(2, pos = 1900-1)
```


Subset of the data for which there are abstracts.

```{r}
absArticles <- allArticles[!is.na(allArticles$Abstract), ]
```

Check distribution of lengths of abstracts
```{r}
par(las = 1)
nCA <- nchar(absArticles$Abstract)
hist(nCA, main = "Distribution of the number of characters in abstracts", 
     xlab = "Number of characters", breaks = seq(0, max(nCA)+50, by = 50), col = cols[3], border = "white", 
     axes = FALSE)
axis(1, pos = 0)
axis(2, pos = 0)
```

Count the number of words
```{r}
nbWordsAbs <- unlist(lapply(strsplit(absArticles$Abstract, " "), length))
hist(nbWordsAbs, main = "Distribution of the number of words in abstracts", 
     xlab = "Number of words", breaks = seq(0, max(nbWordsAbs)+50, by = 10), col = cols[3], border = "white", 
     axes = FALSE)
axis(1, pos = 0)
axis(2, pos = 0)
```


There are `r nrow(absArticles)` such items. 

# Search for words

Function to find a specific word in an abstract
```{r}
findWord <- function(word, line){
  grepl(pattern = word, x = absArticles[line, "Abstract"], ignore.case = TRUE)
}
```

Function to count occurrences of a specific word   
**NB: the proportions are calculated among articles for which there are abstracts**

```{r}
countWord <- function(word, ...){
  v <- unlist(lapply(as.list(1:nrow(absArticles)), function(i) findWord(word, i)))
  
  word.byY <- aggregate(v, by = list(absArticles$PublicationYear), FUN = mean, na.rm = TRUE)
  
  par(las = 1)
  plot(range(word.byY$Group.1), range(word.byY$x), # Initialize the plot
       ylim = c(0, 1), 
       xlab = "Publication Year", ylab = paste0("Proportion of abstracts containing the word `", word, "`"), 
       main = word,
       axes = FALSE,
       type = "n")
  #rect(1900-1, 0, 2020, 1, col = gray(0.9), border = gray(0, 0))
  
  #for(i in seq(0.1, 1, by = 0.1)){
  #  lines(c(1900-1, 2020), rep(i, 2), col = gray(0.95), lwd = 1)
  #}
  
  axis(1, pos = 0)
  axis(2, pos = 1900-1)
  
  # Add the points
  # (doing this so that the points are above the x axis...)
  points(word.byY$Group.1, word.byY$x, pch = 16, col = maincol) 
  
  v
}
```



## Model 

```{r}
absArticles$wordModel <- countWord("model")
```

```{r}
xx <- countWord("model organism")
```

```{r}
xx <- countWord("model system")
```

## Theor*

```{r}
absArticles$wordTheor <- countWord("theor")
```

### Theory

```{r}
xx <- countWord("theory")
```

### Theoretical

```{r}
xx <- countWord("theoretical")
```

## Conceptual

```{r}
absArticles$wordConceptual <- countWord("conceptual")
```

## Equation

```{r}
absArticles$wordEquation <- countWord("equation")
```

## Analy*

```{r}
absArticles$wordAnaly <- countWord("analy")
```

```{r}
xx <- countWord("analytic")
```

```{r}
xx <- countWord("analyze")
```

```{r}
xx <- countWord("analyse")
```

```{r}
xx <- countWord("analytical")
xx <- countWord("analytical solution")
```

## Simulat*

```{r}
absArticles$wordSimulat <- countWord("simulat")
```

## Quantitative

```{r}
absArticles$wordQuantitative <- countWord("quantitative")
```

## Experiment

```{r}
absArticles$wordExperiment <- countWord("experiment")
```

# Citations

```{r}
makeTransparent<-function(someColor, alpha=100)
{
  newColor<-col2rgb(someColor)
  apply(newColor, 2, function(curcoldata){rgb(red=curcoldata[1], green=curcoldata[2],
    blue=curcoldata[3],alpha=alpha, maxColorValue=255)})
}
```

Average per year and per type

```{r fig.height = 7}
par(las = 1)
TC.byY.type <- aggregate(absArticles$TimesCitedWOS, by = list(absArticles$PublicationYear, absArticles$wordModel), FUN = mean)
colsModel <- rep(cols[3], nrow(TC.byY.type))
colsModel[!TC.byY.type$Group.2] <- cols[4]

colsModelTrp <- makeTransparent(colsModel, 200)

plot(TC.byY.type$Group.1, TC.byY.type$x, col = colsModelTrp, xlim = c(1940, 2020), pch = 16, 
     xlab = "Year", ylab = "Average number of citations")
legend(2000, 600, col = c(cols[3], cols[4]), legend = c("Model", "no Model"), pch = 16)
```


```{r}
par(las = 1)
colsModel <- rep(cols[3], nrow(absArticles))
colsModel[!absArticles$wordModel] <- cols[4]
colsModelTrp <- makeTransparent(colsModel, 150)
plot(absArticles$PublicationYear, absArticles$TimesCitedWOS+1, col = colsModelTrp, log = "y", xlim = c(1955, 2020), pch = 16, xlab = "Year", ylab = "Number of citations (+1), log scale")


plot(absArticles$PublicationYear, absArticles$TimesCitedWOS, col = colsModelTrp, log = "", xlim = c(1955, 2020), pch = 16, xlab = "Year", ylab = "Number of citations")
legend(2010, 5000, col = c(cols[3], cols[4]), legend = c("Model", "no Model"), pch = 16)
```

Citations, all items (including those without abstracts)

```{r}
par(las = 1)
plot(allArticles$PublicationYear, allArticles$TimesCitedWOS, col = gray(0.7), log = "", xlim = c(1955, 2020), pch = 16, xlab = "Year", ylab = "Number of citations")
```

Show articles with more than 2000 citations

```{r}
xx <- allArticles[allArticles$TimesCitedWOS > 2000, c("TimesCitedWOS", "Authors", "Title")]
ixx <- sort(xx$TimesCitedWOS, index.return = TRUE, decreasing = TRUE)$ix
xx[ixx, ]
```


